%  Mathpartir --- Math Paragraph for Typesetting Inference Rules
%
%  Copyright (C) 2001, 2002, 2003, 2005, 2015, 2020, 2022, 2024
%  by Didier Remy <didier.remy(at)inria(dot)fr>
%
%  Author         : Didier Remy
%  Version        : 1.5.3
%  Bug Reports    : to author
%  Web Site       : http://cambium.inria.fr/~remy/latex/
%
%  Mathpartir is free software; you can redistribute it and/or modify
%  it under the terms of the GNU General Public License as published by
%  the Free Software Foundation; either version 2, or (at your option)
%  any later version.
%
%  Mathpartir is distributed in the hope that it will be useful,
%  but WITHOUT ANY WARRANTY; without even the implied warranty of
%  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%  GNU General Public License for more details
%  (http://pauillac.inria.fr/~remy/license/GPL).
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  File mathpartir.tex (Documentation)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass {article}

\usepackage {minepartir}
\usepackage {listings}
\usepackage {array}
\usepackage {url}
\newif \ifhevea
%HEVEA \heveatrue
\ifhevea
\usepackage {hevea}
\fi
\def \L#1{\lower 0.4ex \hbox {#1}}
\def \R#1{\raise 0.4ex \hbox {#1}}
\def \hevea {H\L{E}\R{V}\L{E}A}
\def \hevea {$\mbox {H}\!_{\mbox {E}}\!\mbox {V}\!_{\mbox {E}}\!\mbox {A}$}
\def \hemph #1{\begingroup \color{blue}#1\endcolor}

\lstset {basicstyle=\tt}
\let \lst \verb

\title {
%HEVEA\begin{tabular}{>{\huge}c}
\textbf{MATH} formulas in \\
\textbf{PAR}agraph mode\\[-1ex]
\ifhevea\else{}\hskip 0em\hbox to 12em {\hrulefill}\\\fi
%HEVEA\hline
\textbf Typesetting \textbf Inference \textbf Rules \\
%HEVEA\end{tabular}
}
\author {Didier R{\'{e}}my}
\date {(Version 1.4.0, last modified 15/02/2020)}


\begin{document}

\maketitle

\begin{abstract}
This package provides macros for displaying lists of formulas that are
typeset in mixed horizontal and vertical modes. The package is two-folded.

The first part is an environment \verb"mathpar" that generalizes the math
display mode to allow several formulas on the same line, and several lines
in the same display. The arrangement of the sequence of formulas into lines
is automatic depending on the line width and on a minimum inter-formula
space and line width alike words in a paragraphs (in centerline mode).  A
typical application is  displaying a set of type inference rules.

The second par is a macro \lst"inferrule" to typeset inference rules
themselves.  Here again, both premises and conclusions are presented as list
of formulas that should be displayed in almost the same way, except that the
width is not fixed in advance; and the inference rule should use no more
width than necessary so that other inference rules are given a chance to
appear on the same line.

Although \verb"mathpar" and \verb"inferrule" look similar in their
specification, and are often used in combination, they are in fact
completely different in their implementations.
\end{abstract}

\ifhevea
 \section {License}

Mathpartir is Copyright (C) 2001, 2002, 2003, 2005, 2015 INRIA.  Mathpartir
has been developed by Didier R{\'{e}}my.  Mathpartir is free software; you
can redistribute it and/or modify it under the terms of the GNU General
Public License as published by the Free Software Foundation; either version
2, or (at your option) any later version.  See the GNU General Public
License for more details (\url{http://pauillac.inria.fr/~remy/license/GPL}).
Mathpartir is distributed in the hope that it will be useful, but without
any warranty.
\fi

\section {The mathpar environment}

The mathpar environment is a ``paragraph mode for formulas''.
It  allows to typeset long list of formulas putting as
many as possible on the same line:
$$
\begin{tabular}{m{0.45\hsize}m{0.45\hsize}}
\begin{lstlisting}{Ocaml}
\begin{mathpar}
A-Formula \and
Longer-Formula \and
And \and The-Last-One
\end{mathpar}
\end{lstlisting}
&
\begin{mathpar}
A-Formula
\and
Longer-Formula
\and
And
\and
The-Last-One
\end{mathpar}
\end{tabular}
$$
Formulas are separated by \verb"\and" (or equivalently by a blank line).
To enforce a vertical break it suffices to replace \verb"\and" by
\verb"\\".

The implementation of \verb"mathpar" entirely relies on the paragraph mode
for text. It starts  a new paragraph, and a math formula within a paragraph,
after adjusting the spacing and penalties for breaks. Then, it simply binds
\verb"\and" to something like \verb"\goodbreak".

Since \texttt{version 1.5.0}, the environment \verb"mathpar"
reduces the vertical skip using \TeX\ variable
\verb"\abovedisplayshortskip" instead of \verb"\abovedisplayskip"
when the first line of the display is short enough not to overlap with the
last line before the
display:
\begin{mathpar}
one

two

three
\\
four

five
\end{mathpar}
This is to be compared with
\begin{mathpar}
one

two

three
\end{mathpar}
which would overlap with the previous line.
This behavior is automatic.
Currently, it can be canceled with
\verb"\shortmathparfalse", in case of a problem with the new behavior.
However, there should be no reason to do so and this options may be removed
in the future.

An alternative environment \texttt{mathparpagebreakable} allows for page
breaks in the middle of the display.  This cancels the previous optimisation
for short displays.


\section {The inferrule macro}

The inferrule macro is designed to typeset inference rules.  It should
only\footnote {Even though the basic version may work in text mode,
we discourage its use in text mode; the star-version cannot be used in
text-mode} be used in math mode (or display math mode).

The basic use of the rule is
\begin{verbatim}
\inferrule
  {one \\ two \\ three \\ or \\ more \\ premisses}
  {and \\ any \\ number \\ of \\ conclusions \\ as \\ well}
\end{verbatim}
This is the rendering on a large page
\def \one {\inferrule
  {one \\ two \\ three \\ or \\ more \\ premises}
  {and \\ any \\ number \\ of \\ conclusions \\ as \\ well}
}
$$
\ifhevea \one
\else
\fbox {\vbox {\advance \hsize by -2\fboxsep \advance \hsize by -2\fboxrule
       \linewidth\hsize
        $$\one$$}}
\fi
$$
However, the same formula on a narrower page will automatically be typeset
like that:
$$
\ifhevea \one
\else
\fbox {\hsize 0.33\hsize \vbox {$$\one$$}}
\fi
$$
An inference rule is mainly composed of a premise and a conclusion.
The premise and the conclusions are both list of formulas where the
elements are separated by \verb"\\".

Note the asymmetry between typesetting of the premises and of
conclusions where lines closer to the center are fit first.

A newline can be forced by adding an empty line \verb"\\\\"

\begin{tabular}{m{0.44\hsize}m{0.44\hsize}}
\begin{lstlisting}{Ocaml}
\inferrule
   {aa \\\\ bb}
   {dd \\ ee \\ ff}
\end{lstlisting}
&
$\inferrule {aa \\\\bb}{dd \\ ee \\ ff}$
\\
\end{tabular}

\subsection {Single rules}

Single rules are the default mode.
Rules are aligned on their fraction bar, as illustrated below:
$$
\inferrule {aa \\ bb}{ee} \hspace {4em} \inferrule {aa \\\\ bb \\ ee}{ee}
$$
If the premise or the conclusion is empty, then the fraction bar is not
typeset and the premise or the conclusion is centered (when both of them
are empty, which does not make sense, we substitute them by question marks
with a warning):
$$
\begin{tabular}{m{0.54\hsize}m{0.3\hsize}}
\begin{lstlisting}{Ocaml}
\inferrule {}{aa} +
\inferrule {aa \\\\ aa}{} +
\inferrule {}{}
\end{lstlisting}
&
$
\inferrule {}{aa} +
\inferrule {aa \\\\ aa}{} +
\inferrule {}{}
$
\\
\end{tabular}
$$
Use use \verb"{ }" instead of \verb"{}" to typeset an axiom. For instance:
$$
\begin{tabular}{m{0.54\hsize}m{0.3\hsize}}
\begin{lstlisting}{Ocaml}
\inferrule { }{aa} +
\inferrule {aa}{ }
\end{lstlisting}
&
%HEVEA\qquad
\mbox {$
\inferrule { }{aa} +
\inferrule {aa}{ }
$}
\\
\end{tabular}
$$
The macro \lst"\inferrule" accepts a label as optional argument, which will
be typeset on the top left corner of the rule:
\par
\begin{tabular}{m{0.54\hsize}m{0.3\hsize}}
\begin{lstlisting}{Ocaml}
\inferrule [Yop]
   {aa \\ bbc \\ cc}
   {dd}
\end{lstlisting}
&
$\inferrule [Yop] {aa \\ bbc \\ cc} {dd}$
\\
\end{tabular}
\par\noindent
See section~\ref {options} for changing typesetting of labels.
A label can also be placed next to the rule directly, since the rule is
centered:
\par
\begin{tabular}{m{0.54\hsize}m{0.3\hsize}}
\begin{lstlisting}{Ocaml}
\inferrule
   {aa \\\\ bbc \\ cc}
   {cc}
\quad(\textsc {Yop})
\end{lstlisting}
&
$\inferrule {aa \\\\ bbc \\ cc}{cc} \quad (\textsc {Yop})$
\\
\end{tabular}

\subsection {Customizing presentation}

By default, lines are centered in inference rules.
However, this can be changed by either \lst"\mprset{flushleft}"
or \lst"\mprset{center}". For instance,

\begin{tabular}{m{0.44\hsize}m{0.44\hsize}}
\begin{lstlisting}{Ocaml}
$$\mprset{flushleft}
  \inferrule
    {a \\ bbb \\\\ ccc \\ dddd}
    {e \\ ff \\ gg}$$
\end{lstlisting}
&
$$%\mprset{flushleft}
%%BEGIN LATEX
\hsize 0.45\hsize
%%END LATEX
%%HEVEA \hspace{4em}
\inferrule {a \\ bb  \\\\
         ccc \\ dddd}{e \\ ff \\ gg}$$
\\
\end{tabular}

\noindent
Note that lines are aligned independently in the premise and the
conclusion, which are both themselves centered. In particular,
left alignment will not affect a single-line premise or conclusion.

\subsection {Customizing rules}

One may wish to change use rules for rewriting rule or implications, etc.
There is a generic way of definition new rules by providing three parts:
a tail, a body, and a head. The rule will then be built by joining
all three components in this order and filling the body with leaders to
extend as much as necessary. Here are examples

\begin{tabular}{m{0.54\hsize}m{0.44\hsize}}
\begin{lstlisting}{Ocaml}
$$\mprset{fraction={===}}
  \inferrule {a \\ bbb} {cc}$$
\end{lstlisting}
&
$$\mprset{fraction={===}}
  \inferrule {a \\ bbb} {cc}$$

\\
\begin{lstlisting}{Ocaml}
$$\mprset
  {fraction={\models=\Rightarrow}}
 \inferrule {a \\ bbb} {cc}$$
\end{lstlisting}
&
$$\mprset
  {fraction={\models=\Rightarrow}}
  \inferrule {a \\ bbb} {cc}$$
\\
\end{tabular}
The height and depth of the \emph{body} are used to adjust vertical space.
One, may ``smash'' the body to reduce the vertical space

\begin{tabular}{m{0.54\hsize}m{0.44\hsize}}
\begin{lstlisting}{Ocaml}
$$\mprset
  {fraction={%
      {\scriptstyle\vdash}%
      {\smash-}%
      {\rightarrow\!\!}%
      }}
 \inferrule {a \\ bbb} {cc}\,\,$$
\end{lstlisting}
&
$$\mprset
  {fraction={{\scriptstyle\vdash}{\smash-}{\rightarrow\!\!}}}
  \inferrule {a \\ bbb} {cc}$$
\\
\begin{lstlisting}{Ocaml}
$$\mprset {fraction={\cdot\cdots\cdot}}
 \inferrule {a \\ bbb} {cc}$$
\end{lstlisting}
&
$$\mprset {fraction={{}{\,\smash\cdot\,}{}}}
  \inferrule {a \\ bbb} {cc}$$
\\
\end{tabular}
Since vertical skip does not take header and footer into account, which is
usually better but sometimes odd, this can be adjusted explicitly:

\begin{tabular}{m{0.54\hsize}m{0.44\hsize}}
\begin{lstlisting}{Ocaml}
$$\mprset
  {fraction={|=/},
   fractionaboveskip=0.6ex,
   fractionbelowskip=0.4ex}
 \inferrule
   {a \\ bbb_{\downarrow}}
   {cc^{\T\uparrow}}$$
\end{lstlisting}
&
$$\mprset
  {fraction={|=/},
   fractionaboveskip=0.6ex,
   fractionbelowskip=0.4ex}
 \inferrule
   {a \\ bbb_{\downarrow}}
   {cc^{\uparrow}}$$
\\
\end{tabular}
Finally, it is also possible to provide its own definition
of fraction by

\begin{tabular}{m{0.54\hsize}m{0.44\hsize}}
\begin{lstlisting}{Ocaml}
\def \Over #1#2{\hbox{$#1 \over #2$}}
$$\mprset{myfraction=\Over}
  \inferrule {a \\ bbb} {cc}$$
\end{lstlisting}
&
\def \Over #1#2{\hbox{$#1 \over #2$}}
$$\mprset{myfraction=\Over}
  \inferrule {a \\ bbb} {cc}$$
\\
\end{tabular}

\paragraph{Customizing the horizontal skip between premises}
(default value is 2em).
\begin{quote}
\begin{lstlisting}{Ocaml}
$$\mprset {sep=6em}
 \inferrule {a \\ bbb} {cc}$$
\end{lstlisting}
$$\mprset {sep=6em}
  \inferrule {a \\ bbb} {cc}$$
\end{quote}
\paragraph{Customizing the vertical space between premises}

(default value is empty).  Notice that leaving it empty and setting vskip to
0em is not quite equivalent as show below between the third and fourth rules
(because the typesetting cannot use the primitive typesetting of
fractions).
\begin{quote}
\begin{lstlisting}{Ocaml}
$$\def\R{\inferrule {aa\\aa\\\\bbb\\bbb} {cc}
  \hspace{3em}}
  \R  \mprset{vskip=0ex}\R  \mprset{vskip=1ex}\R$$
\end{lstlisting}
$$\def \R{\inferrule {aa \\ aa  \\\\ bbb \\ bbb} {cc}\hspace{3em}}
  \R  \mprset{vskip={}}
  \R  \mprset{vskip=0ex}\R  \mprset{vskip=1ex}\R$$
\end{quote}



\subsection {Tabulars in inference rules\label {sec/braces}}


Although you probably do not want to do that, you may still use tabular,
array or minipage environments inside inference rules, but between braces,
as follows:
$$
\begin{tabular}{m{0.50\hsize}m{0.50\hsize}}
\begin{lstlisting}{Ocaml}
\infer [Tabular-Rule]
{some \\ math \\ and \\
 {\begin{tabular}[b]{|l|r|}
 \hline Ugly & and
  \\[1ex]\hline
  table & text
  \\\hline
 \end{tabular}} \\
 {\begin{minipage}[b]{6em}
  Do you really wish
  to do that?
  \end{minipage}} \\
}
{some \\ conclusions}
\end{lstlisting}
&
\infer [Tabular-Rule]
{some \\ math \\ and \\
 {\begin{tabular}[b]{|l|r|}
 \hline Ugly & and
  \\[1ex]\hline
  table & text
  \\\hline
 \end{tabular}} \\
 {\begin{minipage}[b]{6em}
  Do you really wish
  to do that?
  \end{minipage}} \\
}
{some \\ conclusions}
\\
\end{tabular}
$$


\subsection {Derivation trees}

To help writing cascades of rules forming a derivation tree, inference rules
can also be aligned on their bottom line. For this, we use the star-version:
$$
\begin{tabular}{m{0.65\hsize}m{0.45\hsize}}
\begin{lstlisting}{Ocaml}
\inferrule*
   {\inferrule* {aa \\ bb}{cc}
    \\ dd}
   {ee}
\end{lstlisting}
&
$
\inferrule*
   {\inferrule* {aa \\ bb}{cc}
    \\ dd}
   {ee}
$
\\
\end{tabular}
$$
The star version can also take an optional argument,
but with a different semantics. The optional argument is parsed by the
\verb"keyval" package, so as to offer a set of record-like options:
$$
\def \arraystretch {1.4}
\begin{tabular}{|>{\tt}c|>{$}c<{$}|p{0.6\hsize}|}
\hline
\bf key & \bf arg & \bf Effect
\\\hline
before & tex & Execute $tex$ before typesetting the rule.
         Useful for instance to change the maximal width of the rule.
\\\hline
width & d &  Set the width of the rule to $d$
\\\hline
narrower & d & Set the width of the rule to $d$ times \verb"\hsize".
\\\hline
lab & \ell & Put label $\ell$ on  the top of the rule as with the
non-start version.
\\\hline
Lab & \ell & same as lab
\\\hline
left & \ell & Put label $\ell$ on the left of the rule
\\\hline
Left & \ell & Idem, but as if label $\ell$ had zero width.
\\\hline
Right & \ell & As \verb"Left", but on  the right of the rule.
\\\hline
right & \ell & As \verb"left",  but on the right of the rule.
\\\hline
leftskip & d & Cheat by (skip negative space) $d$ on the left side.
\\\hline
rightskip & d & Cheat by $d$ on the right side of the rule.
\\\hline
vdots & d & Raise the rule by $d$ and insert vertical dots.
\\\hline
\end{tabular}
$$

We remind at the end the global options that we've seen above that can
also be set locally in derivation trees:
\begin{mathpar}
\def \arraystretch {1.4}
\begin{tabular}{|>{\tt}c|>{$}c<{$}|p{0.6\hsize}|}
\hline
\\\hline\hline
sep & d & Set the separation between premises and conclusions to $s$.
\\\hline
flushleft & - & flush premises to the left hand side
\\\hline
center & - & center premises on each line.
\\\hline
rewrite & d &
\\\hline
myfraction & tex & set fraction to $tex$ command
\\\hline
fraction & lmr & set fraction pattern to $lm...mr$ with leaders.
\\\hline
vskip & d & Set the vertical skip between premises and conclusions to $h$.
\\\hline
vcenter && Make the rule centered around the fraction line as the non-star
version
\\\hline
\end{tabular}
\end{mathpar}

Here is an example of a complex derivation:
$$
\inferrule* [left=Total,rightstyle=\em,right={(when $n > 0$)}]
  {\inferrule* [Left=Foo]
     {\inferrule* [rightstyle={\bf},Right=Bar,vskip=1ex,
                   leftskip=2em,vdots=1.5em,rightskip=2em]
         {a \\ a \\\\ bb \\ cc \\ dd}
         {ee}
       \\ ff \\ gg}
     {hh}
  \\
  \inferrule* [lab=XX,sep=4em]{uu \\ vv}{ww}}
  {(1)}
$$
and its code
\begin{lstlisting}{Ocaml}
\inferrule*
  [left=Total,rightstyle=\em,right={(when $n > 0$)}]
  {\inferrule* [Left=Foo]
     {\inferrule*
        [Right=Bar,rightstyle=\bf,
         leftskip=2em,rightskip=vdots=1.5em]
         {a \\ a \\\\ bb \\ cc \\ dd}
         {ee}
       \\ ff \\ gg}
     {hh}
  \\
  \inferrule* [lab=XX]{uu \\ vv}{ww}}
  {(1)}
\end{lstlisting}


\subsection {Label styles}

\label {options}

The package uses
\verb"\DefTirNameStyle",
\verb"\LabTirNameStyle",
\verb"\LeftTirNameStyle",
and \verb"\RightTirNameStyle"
to typeset labels introduced with the default option,
\verb"Lab-",
\verb"Left-", or
\verb"Right-", respectively (or their uncapitalized variants).
This can safely be redefined by the user.
\verb"\DefTirName" is normally used for defining
occurrences ({\em i.e.} in rule \lst"\inferrule") while the three other forms
are used for referencing names ({\em i.e.} in the star-version).  The styles
can also be redefined using labeled-arguments of the star-version of
{\tt\string\inferrule} as described in table below.

Instead of just changing the style, the whole
typesetting of labels may be changed by redefining the commands
\verb"\DefTirName",
\verb"\LabTirName",
\verb"\LeftTirName",
and \verb"\RightTirName", each of which receives the label to be typeset as
argument.

Notice, that if the package \verb"hyperref" is loader, one can automatically
attach hypertarget to rule definitions names of rules defined with the
non-star version of \verb"\inferrule" and refers to them by
\verb"\RefTirName" defined as follows:
\begin{verbatim}
\renewcommand{\DefTirName}[1]{\hypertarget{#1}{\TirName {#1}}}
\newcommand{\RefTirName}[1]{\hyperlink{#1}{\TirName {#1}}}
\end{verbatim}

\begin{mathpar}
\def \arraystretch {1.4}
\begin{tabular}{|>{\tt}c|>{$}c<{$}|p{0.6\hsize}|}
\hline
\bf key & \bf arg & \bf Effect
\\\hline\hline
style & tex & set the default style for labels to $tex$
\\\hline
leftstyle & tex & idem for  labels
\\\hline
rightstyle & tex& idem for right  labels
\\\hline
\end{tabular}
\end{mathpar}


\subsection {Star \emph{v.s.} non-star version}

The package also defines \verb"\infer" as a shortcut for \verb"\inferrule"
but only if it is not previously defined.

There are two differences between the plain and star versions of
\verb"\inferrule".
The plain version centers the rule on the fraction line, while the
star one centers the rule on the last conclusion, so as to be used in
derivation trees.

Another difference is that the optional argument of the plain version is a
label to always be placed on top of the rule, while the $\ast$-version takes
a record of arguments.  Hence, it can be parameterized in many more ways.

One may recover the plain version from the star version by passing the
extra argument \texttt{vcenter} as illustrated below (the base line is
aligned with the dotted line):
\begin{mathpar}
\cdots\cdots
\cdots\cdots
\inferrule*
      {aa  aa \\\\ aa \\ bb}{cc \\ cc \\\\ dd}
\cdots\cdots
\cdots\cdots
\inferrule* [vcenter]
      {aa  aa \\\\ aa \\ bb}{cc \\ cc \\\\ dd}
\cdots\cdots
\cdots\cdots
\end{mathpar}
This is convenient, for instance to typeset rules with side conditions
and keep them attached to the rule:
\begin{mathpar}
\def \RightTirName #1{\rm\hbox {\hskip 1ex (#1)}}
\inferrule*[lab=Pos,right={if $n>0$}]
        {aa \\  aa}
        {cc}

\inferrule*[lab=Neg,right={if $n<0$}]
        {aa \\  aa}
        {cc}
\end{mathpar}
Or differently,
\begin{mathpar}
\def \LabTirName #1{\hbox {(#1)}}
\def \LeftTirName #1{\textsc{#1}}
\inferrule*[Left=Pos,lab={if $n>0$}]
        {aaa \\  aaa}
        {cc}

\inferrule*[Left=Neg,lab={if $n<0$}]
        {aaa \\  aaa}
        {cc}
\end{mathpar}

\subsection{Triple rules}

There is also a version \texttt{\string\triplerule} originally design to print
Hoare triples as rules, which will be rendered as follows, horizontally
aligning rules at their bottom horizontal rule line.
\begin{mathpar}
{\ifhevea\else\hsize 8em\fi
\triplerule[Hoare]
    {ppppp \\  pppp \\ pppp}
    {mmm \\ mmm \\ mm}
    {ccccc \\ cccc}
}

{\ifhevea\else\hsize 8em\fi
\triplerule[Rule]
    {ppppp \\  pppp \\ pppp}
    {mmm \\ mmm}
    {ccccc \\
       {\left(\begin{array}{c}
         top \cr
         bot
         \end{array}\right)}
    }
}
\end{mathpar}
The first rule typeset as follows (in an environment where
\texttt{\string\hsize} is 8em to folow its typesetting on several lines).
\begin{lstlisting}{Ocaml}
\triplerule[Hoare]
    {ppppp \\  pppp \\ pppp}
    {mmm \\ mmm \\ mm}
    {ccccc \\ cccc}
\end{lstlisting}
Here is the code for the second rule
\begin{lstlisting}{Ocaml}
\triplerule[Rule]
     {ppppp \\  pppp \\ pppp}
     {mmm \\ mmm}
     {ccccc \\
        {\left(\begin{array}{c}
          top \cr
          bot
          \end{array}\right)}
     }
}
\end{lstlisting}
Notice that the array environment is surrounded by braces, as
explaned in section \S\ref {sec/braces}.

\subsection {Implementation}

The main macro in the implementation of inference rules is the one that
either premises and conclusions.  The macros uses two box-registers one
\verb"hbox" for typesetting each line and one \verb"vbox" for collecting
lines. The premise appears as a list with
\verb"\\" as separator. Each element is considered in turn typeset in a
\verb"hbox" in display math mode. Its width is compare to the space left on
the current line. If the box would not fit, the current horizontal line is
transferred to the vertical box and emptied. Then, the current formula can
safely be added to the horizontal line (if it does not fit, nothing can be
done). When moved to the vertical list, lines are aligned on their center
(as if their left-part was a left overlapped). At the end the vbox is
readjusted on the right.

This description works for conclusions.  For premises, the elements used to
be processes in reverse order and the vertical list is simply built upside
down, which was annoying when using counters in the premises.  This has
recently been fixed, with a quite different implementation.  In case of
problem, the old behavoir can be recovered by calling
\verb"\MprRecoverOlderVerPremise".

For example,
\begin{lstlisting}{Ocaml}
\newcounter \clab
\newcommand {\lab}[1]
  {\stepcounter{\clab}#1~(\theclab}){}}
\infer [New]
   {\lab A \\\\ \lab B \\ \lab C}
   {\lab D\\ \lab E \\ \lab F}
\end{lstlisting}
produces:
\begin{mathpar}
\newcounter{clab}
\newcommand {\lab}[1]
  {\stepcounter{clab}#1~(\theclab)}
\newcommand {\foo}[1]{
\infer [#1 behavior]
   {\lab A \\\\ \lab B \\ \lab C}
   {\lab D\\ \lab E \\ \lab F}
}
\foo{New}

\MprRecoverOlderVersion
\setcounter{clab}{0}\foo{Old}
\end{mathpar}

\section {Other Options for the {\tt mathpar} environment}

The vertical space in \verb"mathpar" is adjusted by
\verb"\MathparLineskip". To restore the normal paragraph parameters in mathpar
mode (for instance for some inner paragraph), use the command
\verb"\MathparNormalpar".
The environment uses \verb"\MathparBindings" to
rebind \verb"\\", \verb"and", and \verb"\par". You can redefine thus command
to change the default bindings or add your own.


\section {Examples}

See the source of this documentation ---the file \lst"mathpartir.tex"---
for full examples.

\section {{\hevea} compatibility}

The package also redefines \verb"\hva" to do nothing in \LaTeX mode in
both \lst"mathpar" environment nor in inference rules

In \hevea mode and \verb"mathpar" environment, \verb"\and" will always
produce a vertical break in mathpar environment; to obtain a horizontal
break, use \verb"\hva \and" instead.
In \hevea mode inference rules, \verb"\\" and \verb"\\\\"
will behaves as in tex.  In addition you may write \verb"\hva\\\\" to
produce the behavior \verb"\\" and \verb"\\\hva" to produce the behavior of
\verb"\\\\".

For instance, by default the following code,
\begin{lstlisting}{Ocaml}
\begin{mathpar}
\inferrule* [Left=Foo]{}{}

\inferrule* [Left=Foo]
   {\inferrule* [Right=Bar,width=8em,
                 leftskip=2em,rightskip=2em,vdots=1.5em]
      {a \\ a \\ bb \\ cc \\ dd}
      {ee}
    \\ ff \\ gg}
   {hh}
\and
\inferrule* [lab=XX]{uu \\ vv}{ww}
\end{mathpar}
\end{lstlisting}
which typesets in {\TeX} as follows,
\begin{mathpar}
\inferrule* [Left=Foo]
   {\inferrule* [Right=Bar,width=8em,
                 leftskip=2em,rightskip=2em,vdots=1.5em]
      {a \\ a \\\hva bb \\ cc \\ dd}
      {ee}
    \\ ff \\ gg}
   {hh}
\and
\inferrule* [lab=XX]{uu \\ vv}{ww}
\end{mathpar}
would appear as follows with the compatible {\hevea} mode:
\begin{mathpar}
\inferrule* [left=Foo]
   {\inferrule* [right=Bar]
      {a \\ a \\ bb \\ cc \\ dd}
      {ee}
    \\ ff \\ gg}
   {hh}
\\
\inferrule* [lab=XX]{uu \\ vv}{ww}
\end{mathpar}
To obtain (almost) the same rendering as in {\TeX}, it could be typed as
\begin{lstlisting}[escapechar=\%]{Ocaml}
  \begin{mathpar}
      \inferrule* [Left=Foo]
      {\inferrule* [Right=Bar,width=8em,
                    leftskip=2em,rightskip=2em,vdots=1.5em]
         {a \\ a \\\hva bb \\ cc \\ dd}
         {ee}
         \\ ff \\ gg}
      {hh}
      \hva \and
  \inferrule* [lab=XX]{uu \\ vv}{ww}
  \end{mathpar}
  \end{lstlisting}
  Actually, it would be typeset and follows with the compatible {\hevea} mode:
  \begin{mathpar}
      \inferrule* [left=Foo]
      {\inferrule* [right=Bar]
         {a \\ a \\\\ bb \\ cc \\ dd}
         {ee}
         \\ ff \\ gg}
      {hh}
      \and
  \inferrule* [lab=XX]{uu \\ vv}{ww}
\end{mathpar}

\end{document}

%% DRAFT
\section {aa}

\inferrule[Lab]{}{}

Checking \texttt{shortskip} around \texttt{display} in the reimplementation of
\texttt{\{mathpar\}}.

\newdimen \bar \bar 3.4pt
\def \rule #1{\hbox to #1\bar{[\dotfill]}}
aaaaa aaaaa aaaaa aaaaa aaaaa aaaaa aaaaa
aaaaa aaaaa aaaaa aaaaa aaaaa aaaaa aaaaa
aaaaa aaaaa aaaaa aaaaa aaaaa aaaaa aaaaa
\\
aaaa
\begin{mathpar}
\rule {20}{\bar 1pt\rule {40}}

\rule {30}

\rule {40}

\rule {40}
\iffalse%\iftrue

\rule {20}

\rule {50}%40
\fi
\end{mathpar}
ccccc
ccccc ccccc ccccc ccccc ccccc ccccc ccccc
ccccc ccccc ccccc ccccc ccccc ccccc ccccc ccccc
ccccc ccccc ccccc ccccc ccccc ccccc ccccc ccccc ccccc
ccccc ccccc ccccc ccccc ccccc ccccc ccccc ccccc ccccc

\end{document}

\section {Testing for potential bugs}

The user could add his own customization:
\makeatletter
\define@key {mpr}{style}[1]{\def\TirNameStyle{#1}What So Ever}
\define@key {mpr}{rightstyle}[1]{\def\RightTirNameStyle{#1}}
\makeatother

includes a spurious "def" which means that when you try to change the style
by using something like
$$
\inferrule*[style=\texttt,lab=Def] A B
$$

\endinput


%% LocalWords: mathpar aa Yop bb dd ee ff cc Ocaml Foo leftskip
%% LocalWords: rightskip vdots gg hh uu vv ww
%% LocalWords: HeVeA escapechar inferrule yop flushleft bbb
%% LocalWords: ccc dddd fractionaboveskip fractionbelowskip myfraction
%% LocalWords: sep vskip Tabulars minipages keyval arg tex lmr vcenter
%% LocalWords: rightstyle uncapitablized leftstyle parameterized Pos aaa
%% LocalWords: hbox vbox mathpartir
